\LaTeXe\ guide for the Journal of Functional Programming

### 2019-05-02 papers/jfp/7-experience.tex
Redrafting in section 8

Based on the thinking in this paper we have extended \Shake to
become \Cloud \Shake, the first cloud-capable build system to support
both early cutoff and monadic dependencies (\S{sec-implementation-cloud}),
and deployed it to GHC's (very substantial) build system, Hadrian~\cite{hadrian}.
In this section we tie these papers together, reflecting on our
experience since these papers.

>>>>

Based on the thinking in this paper we have extended \Shake to
become \Cloud \Shake, the first cloud-capable build system to support
both early cutoff and monadic dependencies (\S\ref{sec-implementation-cloud}),
and used it to implement GHC's (very substantial) build system, Hadrian~\cite{hadrian}.
In this section we tie these papers together, reflecting on our
experience of turning theory into practice.

### 2019-05-02 papers/jfp/8-engineering.tex
Redrafting in section 8

\item \textbf{Non-determinism}: Some tasks are \emph{non-deterministic},
      producing any result from a possible set. As an example, GHC programs
      compiled using parallelism can change the order in which unique variables
      are obtained from the supply, producing different but semantically
      identical results.

\item \textbf{Volatility}: Some tasks are defined to change in every build, e.g.
      \Excel provides a~``function'' \cmd{RANDBETWEEN} producing a fresh random
      number in a specified range on each recalculation. Similarly, build
      systems like \Make and \Shake provide volatile \emph{phony rules}. The
      main difference from non-deterministic tasks is that volatile tasks cannot
      be cached. They are best modelled as depending on a special key
      \cmd{RealWorld} whose value is changed in every build.
\end{itemize}

Interestingly, there is significant interplay between all three sources of
impurity. Systems like \Bazel use various sandboxing techniques to guard against
missing dependencies, but none are likely to capture all dependencies right down
to the CPU model and microcode version. Tasks that do have untracked
dependencies can be marked as volatile, a technique \Excel takes with the
\cmd{INDIRECT} function, removing the untracked dependency at the cost of
minimality.

@@

\noindent
The build systems that have precise self-tracking are all ones which use a
\emph{non-embedded domain specific language} to describe build tasks. Those
which use a full programming language, e.g. \Shake, are faced with the challenge
of implementing equality on arbitrary task functions. For such build systems,
the pessimistic assumption that any change to the build system potentially
changes any build task can often be used~--~the classic example being a makefile
depending on itself.

>>>>

\item \textbf{Non-determinism}: Some tasks are \emph{non-deterministic},
      producing any result from a possible set. As an example, GHC programs
      compiled using parallelism can change the order in which unique variables
      are obtained from the name supply, producing different but semantically
      identical results.

\item \textbf{Volatility}: Some tasks are defined to change in every build, e.g.
      \Excel provides a~``function'' \cmd{RANDBETWEEN} producing a fresh random
      number in a specified range on each recalculation. Similarly, build
      systems like \Make and \Shake provide volatile \emph{phony rules}. The
      main difference from non-deterministic tasks is that volatile tasks cannot
      be cached. They are readily modelled as depending on a special key
      \cmd{RealWorld} whose value is changed in every build.
\end{itemize}

Interestingly, there is significant interplay between all three sources of
impurity. Systems like \Bazel use various sandboxing techniques to guard against
missing dependencies, but none are likely to capture all dependencies right down
to the CPU model and microcode version. Tasks that do have untracked
dependencies can be marked as volatile, a technique \Excel takes with the
\cmd{INDIRECT} function, fixing the untracked dependency at the cost of
minimality.

@@

\noindent
The build systems that have precise self-tracking are all ones which
use a \emph{non-embedded domain specific language} to describe build
tasks; that is, a task is described by a data structure (or syntax
tree), and tasks can be compared for equality by comparing those data
structures. Those which use a full programming language, e.g. \Shake,
are faced with the challenge of implementing equality on arbitrary
tasks -- and a task is just a function.  For such build systems, the
only safe approach is to assume (pessimistically) that any
change to the build system potentially changes any build task ~--~ the
classic example being a makefile depending on itself.
